Eco-evolutionary dynamics of finite populations from first principles

### 2023-03-16 ind_files/06_nontechnical_discussion.tex
Minor phrasing changes

Indeed, the most important takeaway from this thesis is that the equations of section \ref{sec_fun_theorems} are very general, since part \ref{part_theory} makes essentially no assumptions other than density dependence, the impossibility of infinite growth starting from finite population size, and the ability to define per-capita birth and death rates, and thus, these birth and death rate function(al)s can in general be almost arbitrarily complicated. Further, as we saw in section \ref{sec_fun_theorems}, the terms of the equations lend themselves to simple biological interpretation. Like the classical Price equation, the utility of the equations of section \ref{sec_fun_theorems} lies not (necessarily) in their solutions for specific models, but instead in their generality and the fact that their terms help us clearly think about the various evolutionary phenomena operating in biological populations~\citep{frank_natural_2012,luque_one_2017, luque_mirror_2021}. The general spirit of this thesis is thus identical to the philosophy behind the general move towards trying to formulate `model-independent' eco-evolutionary theory, a move that has recently been (rapidly) gaining popularity in the literature~\citep{grafen_formal_2014, queller_fundamental_2017, lion_theoretical_2018, allen_mathematical_2019, rice_universal_2020, week_white_2021, wickman_theoretical_2022, kuosmanen_turnover_2022, mazzolini_universality_2022,lion_extending_2023,allen_natural_2023}. However, there are some biologically important factors that I have neglected in my formalism. My framework works with unstructured populations, neglecting any population structure due to groups, age, class, sex, developmental stage, or space\footnote{Of course,  since position in continuous space is just a special case of a quantitative trait, we have technically incorporated space in a very limited sense: The formalism of chapter \ref{chap_infD_processes} can equally well describe a finite population of exactly identical individuals moving through one-dimensional space, and so can technically model mechanisms like range expansion of clonal populations. But this is only a technicality and such populations are likely somewhat `evolutionarily boring', since all individuals in the model must always be exactly identical in all aspects other than spatial location.}, all of which can lead to very complex and often surprising dynamics. Explicitly incorporating features such as population structure, sex, and space from first principles in an analytically tractable framework is a formidable task that will likely need innovative new mathematical and biological arguments, and presents a fantastic opportunity for future studies.

>>>>

Indeed, the most important takeaway from this thesis is that the equations of section \ref{sec_fun_theorems} are very general, since part \ref{part_theory} makes essentially no assumptions other than density dependence, the impossibility of infinite growth starting from finite population size, and the ability to define per-capita birth and death rates, and thus, these birth and death rate function(al)s can in general be almost arbitrarily complicated. Further, as we saw in section \ref{sec_fun_theorems}, the terms of the equations lend themselves to simple biological interpretation. Like the classical Price equation, the utility of the equations of section \ref{sec_fun_theorems} lies not (necessarily) in their solutions for specific models, but instead in their generality and the fact that their terms help us clearly think about the various evolutionary phenomena operating in biological populations~\citep{frank_natural_2012,luque_one_2017, luque_mirror_2021}. The general spirit of this thesis is thus identical to the philosophy behind the general move towards trying to formulate `model-independent' eco-evolutionary theory, a move that has recently been (rapidly) gaining popularity in the literature~\citep{grafen_formal_2014, queller_fundamental_2017, lion_theoretical_2018, allen_mathematical_2019, rice_universal_2020, week_white_2021, wickman_theoretical_2022, kuosmanen_turnover_2022, mazzolini_universality_2022,lion_extending_2023,allen_natural_2023}. However, there are some biologically important factors that I have neglected in my formalism. My framework works with unstructured populations, neglecting any population structure due to groups, age, class, sex, developmental stage, or space\footnote{Of course,  since position in continuous space is just a special case of a quantitative trait, we have technically incorporated space in a very limited sense: The formalism of chapter \ref{chap_infD_processes} can equally well describe a finite population of exactly identical individuals moving through one-dimensional space, and so can technically describe phenomena like range expansion of clonal populations. However, I am only noting this as a technicality - Such populations are likely somewhat `evolutionarily boring', since all individuals in the model must always be exactly identical in all aspects other than spatial location.}, all of which can lead to very complex and often surprising dynamics. Explicitly incorporating features such as population structure, sex, and space from first principles in an analytically tractable framework is a formidable task that will likely need innovative new mathematical and biological arguments, and presents a fantastic opportunity for future studies.
